sensi_an_rotations1. Start of sensitivity analysis for rotations.

### 2018-05-09 inst/documents/paper3/paper3_rotations.Rmd
pap3

*todo describe these example runs first, before talking about overall results*

@@

@@

@@

Thus to aid the comparison of rotations and sequences model runs can be broadly classified into three groups :

When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold (Fig 4.1).

*todo : maybe I should just leave the 'succeeded' runs in here which would change these %s slightly*
With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold. In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2).

Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 4.3).

With both costs and dispersal the benefits of rotations are similar to when runs have costs alone (Fig 4.4).

@@

Note however that this combination of recessive resistance and high dominance costs are exactly the features most desirable in any insecticide irrespective of how it is deployed. We discuss fitness below but note here that it is very difficult to maintain insecticide concentration at the level required to make resistance recessive. Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant; see Figure 1 of <cite Levick et al> and our more extensive discussion in <cite South et al WoS manuscript).

@@

## TODO
*try to come up with brief name for model output to make it easier to refer to it later*
*add caveat about polygenic resistance*
*The caption to each graph needs to have the list of parameter values used to generate them*
*work on Ians fig Y to help explanation*

@@

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia. The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence reach the endpoint, resistance frequency > 0.5 to all insecicides, at the same time. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and remains at the constant level while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. All other inputs are kept constant between the rotation and the sequence."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$plot <-     FALSE

@@

```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there are resistance fitness costs. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation lasts longer than the sequence. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step but in contrast to the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines  while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0.05 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$plot <-     FALSE

@@

```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3. Comparing the increase in resistance frequencies over time for a rotation and a sequence with dispersal from untreated refugia. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence last the same time. In both cases the resistance frequencies in the treated areas (red lines) increase when an insecticide is in use. Resistance frequencies in the untreated areas (blue lines) increase after those in the treated areas as a result of dispersal from the treated areas. When an insecticide stops being used the resistance frequency in the treated area declines due to dispersal from the untreated area. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and consistent with the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 0.5
linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$plot <-     FALSE

@@

```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 4. Effect of cost and dispersal on the difference between rotations and sequences from model runs. The y axis shows the number of model runs for the different caegories of difference between rotations and sequences. 4.1 when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results. 4.2, 4.4 when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences. For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences. 4.3 when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}

  # remove runs where thresholds not reached for seq & rot
  # with cost_mig ~ 65% of runs don't reach thresholds
  # could get this func to output perecnt not reaching
  rem_thresh_not_reached <- function(x) {x %>% filter(gens_seq != max_gen | gens_rot != max_gen)}

  df_nocost_nomig <- rem_thresh_not_reached(df_nocost_nomig)
  df_cost <- rem_thresh_not_reached(df_cost) #6.6
  df_mig <- rem_thresh_not_reached(df_mig) #0.05
  df_cost_mig <- rem_thresh_not_reached(df_cost_mig) #6.9

@@

  custom_bins <- c(-20,-1,1,20,100,450)
  df_nocost_nomig$rot_adv_bins <- cut(df_nocost_nomig$pcent_rot_adv, custom_bins)
  df_cost$rot_adv_bins <- cut(df_cost$pcent_rot_adv, custom_bins)
  df_mig$rot_adv_bins <- cut(df_mig$pcent_rot_adv, custom_bins)
  df_cost_mig$rot_adv_bins <- cut(df_cost_mig$pcent_rot_adv, custom_bins)

@@

             geom_bar() +
             #geom_histogram(binwidth=10) +
             #xlim(-20,400) +
             xlab('') +
             ylab('count') +
             ggtitle('2. with costs') +
             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
             #scale_x_continuous(breaks=c(-10,-1,0,1,20,100,400)) +
             scale_x_discrete( limits=bin_labs,
                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
             theme_minimal() +
             theme(panel.grid.minor.y=element_blank(),
                   axis.text.x=element_blank())

@@

             geom_bar() +
             xlab('Percent benefit of rotation over sequence') +
             ylab('count') +
             ggtitle('4. with costs and dispersal') +
             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
             theme_minimal() +
             scale_x_discrete( limits=bin_labs,
                               labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+       
             theme(panel.grid.minor.y=element_blank())

@@

```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 5. Example of a model run where there is a substantial benefit of rotations over a sequence. For the sequence resistance thresholds are reached for all insecticides in a little over 100 generations or 10 years. For the rotation resistance levels for all insecticides are maintained well below resistance thresholds for the length of the simulation (500 generations or 50 years)."}
# ? are these run ids before or after the runs that don't reach thresholds have been removed ?
# seems OK that they are after other runs removed
#scenarios_to_plot <- c(624,2167,602,4028,5771)
scenarios_to_plot <- c(2167)

>>>>

@@

*todo continue describing the example runs first, before talking about overall results*

@@

*todo : change the %s because 'succeeded' runs now left in*
When there were no costs or dispersal there was no difference between rotations and sequences in the number of generations below the resistance threshold (Fig 4.1). With fitness costs included 67% of model runs, similar to those in the previous section, gave no difference between rotations and sequences in the number of generations below the resistance threshold. In contrast, 7% of model runs generated an advantage of greater than 20% for rotations over sequences (Fig 4.2). Dispersal from untreated refugia gives smaller benefits of rotations over sequences and in fewer runs (Fig 4.3). With both costs and dispersal the benefits of rotations are similar to when runs have costs alone (Fig 4.4).

@@

Thus to aid the comparison of rotations and sequences model runs can be broadly classified into three groups :
1. both strategies succeeded : resistance thresholds for all insecticides not reached within 500 generations
2. rotations only succeeded
3. neither strategies succeeded

*todo show a histogram plot of the counts of model runs in each category, across all startegies and scenarios*
*decide whether to put before or after fig4*

@@

Note however that this combination of recessive resistance and high dominance costs are exactly the features most desirable in any insecticide irrespective of how it is deployed. We discuss fitness below but note here that it is very difficult to maintain insecticide concentration at the level required to make resistance recessive. Concentrations decay post-application in most deployment strategies and this cause resistance to gradually change from being recessive to being dominant; see Figure 1 of [@Levick2017] and our more extensive discussion in [cite South et al WoS manuscript].

@@

## TODO
*add to methods description of how fitness determined in model, move & modify fig 1*
*try to come up with brief name for model output to make it easier to refer to it later*
*add caveat about polygenic resistance*
*The caption to each graph needs to have the list of parameter values used to generate them*
*work on Ians fig Y to help explanation*

@@

```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 1. "}

#temp moved 
#Comparing the increase in resistance frequencies over time for a rotation and a sequence when there no resistance fitness costs or dispersal from untreated refugia. The upper plot shows a rotation with a regular interval of 10 generations and the lower plot a sequence in which the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence reach the endpoint, with all insecticides having a resistance frequency above 0.5, at the same time. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and that insecticide cannot be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and remains at the constant level while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1.

@@

linputs$n_insecticides <- 3
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

@@

```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 2. Comparing the increase in resistance frequencies over time for a rotation and a sequence when there are resistance fitness costs. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation lasts longer than the sequence. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step but in contrast to the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines  while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0.05, dominance of cost=1."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0.05 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

@@

```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 3. Comparing the increase in resistance frequencies over time for a rotation and a sequence with dispersal from untreated refugia. The upper plot shows a rotation with a regular interval and the lower plot a sequence when the insecticide is changed when a resistance threshold of 0.5 is reached. In this example the rotation and the sequence last the same time. In both cases the resistance frequencies in the treated areas (red lines) increase when an insecticide is in use. Resistance frequencies in the untreated areas (blue lines) increase after those in the treated areas as a result of dispersal from the treated areas. When an insecticide stops being used the resistance frequency in the treated area declines due to dispersal from the untreated area. For the sequence the resistance frequency for each insecticide increases to reach the threshold in one step and consistent with the previous figure resistance frequencies decline when the insecticide is not in use allowing it to be used again. For the rotation the resistance frequency for each insecticide increases in shorter steps while it is in use and similarly declines while not. There are three insecticides, the resistance frequency for each is shown in the sub-plots, and the shaded boxes within each sub-plot show when that insecticide was in use. The simulation is stopped when there are no remaining insecticides for which the resistance frequency is below the threshold. All other inputs are kept constant between the rotation and the sequence. Effectiveness=0.6, exposure=0.6, male exposure proportion=0.5, resistance restoration=0.5, dominance of resistance=1, cost=0, coverage=0.5, dispersal=0.1."}

@@

linputs$n_insecticides <- 3
linputs$coverage <- 0.5
linputs$migration <- 0.1
linputs$cost <-     0 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1
linputs$rr <-       0.5
linputs$male_expo_prop <-  0.5
linputs$plot <-     FALSE

@@

```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='center', fig.cap="Figure 4. Effect of cost and dispersal on the difference between rotations and sequences from model runs. The y axis shows the number of model runs for the different caegories of difference between rotations and sequences. 4.1 when there are no resistance costs or dispersal from untreated refugia, rotations and sequences produce the same results. 4.2, 4.4 when costs are added there are infrequent model runs where rotations last longer below resistance thresholds than sequences. For between 6 and 7% of runs including costs, rotations last more than 20% longer than sequences. 4.3 when dispersal from untreated refugia is included without costs there are even less frequent runs where rotations last longer than sequences (in just 4 out of 10000 runs rotations lasted greater than 20% longer than sequences)."}

  # remove runs where thresholds not reached for seq & rot
  # with cost_mig ~ 65% of runs don't reach thresholds
  # could get this func to output percent not reaching
  # now don't remove the runs where thresholds not reached
  # rem_thresh_not_reached <- function(x) {x %>% filter(gens_seq != max_gen | gens_rot != max_gen)}
  # 
  # df_nocost_nomig <- rem_thresh_not_reached(df_nocost_nomig)
  # df_cost <- rem_thresh_not_reached(df_cost) #6.6
  # df_mig <- rem_thresh_not_reached(df_mig) #0.05
  # df_cost_mig <- rem_thresh_not_reached(df_cost_mig) #6.9

@@

  #custom_bins <- c(-20,-1,1,20,100,450)
  custom_bins <- c(-20,-10,-1,1,10,20,100,450)
  df_nocost_nomig$rot_adv_bins <- cut(df_nocost_nomig$pcent_rot_adv, custom_bins)
  df_cost$rot_adv_bins <- cut(df_cost$pcent_rot_adv, custom_bins)
  df_mig$rot_adv_bins <- cut(df_mig$pcent_rot_adv, custom_bins)
  df_cost_mig$rot_adv_bins <- cut(df_cost_mig$pcent_rot_adv, custom_bins)

@@

             geom_bar() +
             #geom_histogram(binwidth=10) +
             #xlim(-20,400) +
             xlab('') +
             ylab('count') +
             ggtitle('2. with costs') +
             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
             #scale_x_continuous(breaks=c(-10,-1,0,1,20,100,400)) +
             scale_x_discrete( limits=bin_labs,
                               #labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+    
             theme_minimal() +
             theme(panel.grid.minor.y=element_blank(),
                   axis.text.x=element_blank())

@@

             geom_bar() +
             xlab('Percent benefit of rotation over sequence') +
             ylab('count') +
             ggtitle('4. with costs and dispersal') +
             #scale_y_sqrt(breaks=c(0,10,20,50,200,500,2000,5000)) + 
             theme_minimal() +
             scale_x_discrete( limits=bin_labs,
                               #labels=c("-20 to -1", "-1 to 1", "1 to 20", "20 to 100", "100 to 500"))+
                               labels=c("-20 to -10", "-10 to -1", "-1 to 1", "1 to 10", "10 to 20", "20 to 100", "100 to 500"))+     
             theme(panel.grid.minor.y=element_blank())

@@

```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7, fig.width=9, fig.show='hold', fig.align='center', fig.cap="Figure 5. Example of a model run where there is a substantial benefit of rotations over a sequence. For the sequence resistance thresholds are reached for all insecticides in a little over 100 generations or 10 years. For the rotation resistance levels for all insecticides are maintained well below resistance thresholds for the length of the simulation (500 generations or 50 years). Effectiveness=0.57, exposure=0.88, male exposure proportion=0.72, resistance restoration=0.68, dominance of resistance=0.04, cost=0.07, dominance of cost=0.59."}
# ? are these run ids before or after the runs that don't reach thresholds have been removed ?
# seems OK that they are after other runs removed
#scenarios_to_plot <- c(624,2167,602,4028,5771)

#now that non thershold runs are not removed :
#which(df_cost$pcent_rot_adv > 200)
#987 1023 1324 1365 1511 2051 2717 3187 3196 3582 3589 3768 4128 4219 4835 4880 6233 6538 6913 8666 9361
scenarios_to_plot <- c(2717)

### 2018-05-09 inst/documents/rotations_notes.Rmd
pap3

*8/5/18*

>>>>

*8/5/18*
Getting violin plots working.

*9/5/18*
Leaving in all runs now

Something weird has happened to fig1 in the paper3 ?? because of a > in fig legend

Also because we are leaving all runs in, the indices of the runs with large rot advantage are now wrong ! I should have recoded a unique id for runs.

ARg, now all plots have stopped appearing in doc and I don't know why ?????
It was a < symbol in the MS text, could tell because the text stopped at that point.
